### 2025-03-18 paper.qmd
Added Information Regarding Additional Test

---
title: "The Tail Risk Premium - Extension of Lottery Preference Premium"
author: "Brayden Boyko, Roan Henderson, Andrew Pettifor"
format:
  html:
    code_folding: hide
    self_contained: true
    toc: false
    dev: "png"
    fig_embed: true
    embed-resources: true
execute:
  echo: false
jupyter:
  kernel: python3
---
## Abstract
The investment strategy centered around the Idiosyncratic Volatility (IVOL) Risk Premium exploits the distinct performance patterns of stocks exhibiting varying levels of firm-specific risk. Unlike systemic risk, which permeates the entire market, idiosyncratic volatility is unique to individual firms and can be effectively diversified within a well-constructed portfolio. Empirical research consistently demonstrates that stocks with low idiosyncratic volatility tend to outperform those with high idiosyncratic volatility, suggesting a negative risk premium for high-IVOL stocks and a positive premium for low-IVOL stocks. This phenomenon has been validated across diverse international markets and controlled for market, size, and book-to-market factors, forming the foundation of this investment strategy.

## Investment Strategy
The core of the strategy involves taking long positions in stocks characterized by low idiosyncratic volatility, which consistently exhibit a positive return premium. These stocks are frequently underpriced relative to their risk profiles, presenting an opportunity for reliable excess returns. Conversely, stocks with high idiosyncratic volatility are systematically mispriced due to behavioral biases such as investor preference for lottery-like payoffs and market frictions that limit arbitrage. By strategically shorting these high-IVOL stocks, investors can exploit their documented pattern of underperformance. This dual approach leverages market inefficiencies to achieve superior risk-adjusted returns.

## Persistence of the Idiosyncratic Risk Premium:
The persistence of the idiosyncratic risk premium is attributed to several factors, including investor behavior, market frictions, and liquidity constraints. High-IVOL stocks attract risk-tolerant investors seeking high returns, but deter conservative investors, creating a premium for stocks with lottery-like features. Market frictions, such as high transaction costs and information asymmetry, further perpetuate this premium by impeding effective arbitrage. Additionally, liquidity constraints necessitate higher compensation for holding assets with high idiosyncratic risk. By comprehensively understanding and exploiting these dynamics, the investment strategy aims to deliver consistent and predictable returns.

## Strategy Details
To effectively harness the Idiosyncratic Volatility (IVOL) Risk Premium, the strategy employs a systematic approach in selecting and managing investments:

### Data Selection:
+ The initial step involves gathering comprehensive data from the WRDS database by Wharton, spanning from 2002 to 2022. This extensive data set includes stock prices, historical returns, and idiosyncratic volatility measures.
+ The portfolio will be rebalanced annually from 2007 to 2022, ensuring a lookback window of five additional years to refine the selection criteria.

### Portfolio Construction
+ Selection Criteria: The strategy identifies stocks based on idiosyncratic volatility and skewness. Stocks exhibiting high idiosyncratic volatility and negative skewness are shortlisted for short positions, while those with low idiosyncratic volatility and positive skewness are chosen for long positions.

+ Long Positions: The strategy takes long positions in stocks with low idiosyncratic volatility, aiming to capitalize on their consistent positive return premium. These stocks are often underpriced relative to their risk profiles, offering reliable opportunities for excess returns.

+ Short Positions: Simultaneously, short positions are established in stocks with high idiosyncratic volatility. These stocks are systematically mispriced due to behavioral biases and market frictions, and they tend to underperform over time. The strategy seeks to profit from their predictable pattern of underperformance.

### Risk Management:
+ Diversification: The portfolio is diversified across multiple stocks to mitigate firm-specific risks. By constructing a well-balanced portfolio, the strategy ensures that the impact of any single stock's volatility is minimized.
+ Rebalancing: Annual rebalancing is conducted to maintain the desired risk exposure and adjust for changes in idiosyncratic volatility and skewness. This regular review allows the strategy to adapt to evolving market conditions and optimize returns.

### Performance Monitoring:
+ The strategy's performance is continuously monitored against relevant benchmarks to ensure alignment with its objectives. Key performance indicators, such as risk-adjusted returns and volatility measures, are tracked to assess the strategy's effectiveness.
+ Adjustments: Based on performance analysis, adjustments may be made to the portfolio composition, selection criteria, or rebalancing frequency to enhance the strategy's overall performance and resilience to market fluctuations.

## Performance Evaluations
A variety of benchmark statistics will be included in our portfolio testing:

+ Annual / Monthly Return Values/Charts For 4 Portfolios
+ Description of Stock Breakdown For Four Portfolios (Skew, IVOL)
+ V1 & V2 Screening

### Four Portfolio Metrics

Four portfolio's were constructed under the assumption of equal weighting dynamics. V1 portfolios are comprised of 100 companies and screen through IVOL. V2 portfolios describe our teams extension of the V1 sceen, which utilizes skew to screen for 30 companies in each portfolio.

#### Monthly Returns

```{python}
from visuals import monthly_ret

monthly_ret()

```

#### Monthly Return Histogram

```{python}
from visuals import monthly_hist

monthly_hist()
```

#### Cumalitve Monthly Returns (Value of $1 Invested)

```{python}
from visuals import monthly_total_return

monthly_total_return()
```

#### Annual Returns (Value of $1 Invested)
```{python}
from visuals import annual_ret
annual_ret()

```

### Screening Variables

The screening variables are broken into four categories, under the asumption of a three year lookback period. Base Long utilizes the lowest 100 IVOL from the dataset each year. In contrast, the Base Short utilizes the highest 100 IVOL from the dataset.

#### V1 Screen: High-Low IVOL

```{python}
from visuals import monthly_ivol

monthly_ivol()
```

#### V2 Screen: Negative-Positive Skew

```{python}
from visuals import portfolio_skews

portfolio_skews()
```

### Portfolio Positions
The portfolios are equal weighted. This provides an accurate representation of both the V1 and V2 screening effectiveness.

#### V2 Portfolio Positions Over Time

```{python}
from visuals import ticker_count

ticker_count()
```

#### Equal Weighted Portfolios (Baseline vs Strategy-Extension) Cumalitve Returns (Value of $1 Invested)

```{python}
from visuals import portfolio_totals

portfolio_totals()
```

#### V2 Strategy Alpha Relative To V1 Baseline

```{python}
from visuals import monthly_alpha

monthly_alpha()

```

### Additional Tests

A series of additional tests were completed to better explain the results found in our data sample period. These tests were conducted in order to improve upon the strategy.

#### Explantion For V2 Short Measure Underperformance

#### Explanation For IVOL Stabalization Post 2008

__Key Takeaways for Post-2008 Market Stability__

+ Systemic risk increased post-2008, making diversification ineffective.
+ Correlations tend to rise before a crisis, suggesting a leading indicator for instability.
+ Investor behavior shifted towards liquidity and safety, aligning with our discussion on idiosyncratic risk vs. systemic risk.
+ Market-wide herding intensified, driving the preference for safe assets and reducing the appeal of high-idiosyncratic-risk stocks.
+ Relative decline in idiosyncratic risk due to rising market and industry volatility.
+ Crisis events (2008, COVID-19) still caused firm-specific volatility spikes but within an environment of elevated systematic risk.
+ Investor preference for stability (CSP, large firms) helped reduce firm-specific risk, supporting our discussion on flight to safety and liquidity-driven volatility shifts.

1. "Temporal Evolution of Financial Market Correlations" (Fenn et al.)

__Summary:__ This study applies Principal Component Analysis (PCA) and Random Matrix Theory (RMT) to analyze evolving correlations in financial markets from 1999–2010. Key findings indicate that post-2008, financial assets across different markets exhibited much stronger correlations. The first principal component accounted for nearly 40% of market variance post-Lehman collapse, suggesting that most assets were moving together. This heightened correlation reduces diversification benefits and increases systemic risk.

__Relevance to Post-2008 Stability:__ This aligns with our discussion on correlation risk, where increased asset co-movement reduces the effectiveness of diversification. It also supports the high systemic risk in a post-2008 world, showing why investor behavior shifted towards liquidity preference and safe-haven assets.

2. "Correlation of Financial Markets in Times of Crisis" (Sandoval & Franca)

__Summary:__ The study uses eigenvalue decomposition of correlation matrices to show that market correlations increase dramatically during crises (1987, 1998, 2001, 2008). Their key finding is that high volatility leads to strong correlations, meaning that during crashes, markets behave as one. The study also finds that average market correlation increases prior to a crisis, suggesting an early warning signal.

__Relevance to Post-2008 Stability:__ This reinforces our point that in crisis conditions, herding behavior dominates. The finding that correlations increase before crises aligns with the liquidity crunch hypothesis, where investors pull out of riskier assets preemptively. The study also supports the behavioral premium idea, where investors chase "lottery stocks" in stable times but rush to liquidity in crises.

3. "Stock Markets Correlation - Before and During the Crisis Analysis" (Moldovan)​"

__Summary:__ This paper specifically analyzes the correlations between New York (Dow Jones), London (FTSE), and Tokyo (Nikkei) before (2003-2006) and during (2007-2010) the crisis. It finds that correlations were significantly stronger during the crisis period, meaning that stocks moved together more in downturns than in uptrends. Regression models showed that FTSE’s influence on Dow and Nikkei increased during the crisis, while Nikkei and Dow became negatively correlated.

__Relevance to Post-2008 Stability:__ This supports our discussion on global financial interdependence post-2008. The increased correlation in downturns confirms that diversification failed when most needed. The negative Dow-Nikkei correlation is an interesting insight, possibly reflecting a flight to safety in Japanese markets during the US downturn.