### 2025-03-31 Animated Graph.qmd
Updated Data

---
title: ""
author: ""
format: 
  html:
    embed-resources: true
    page-layout: full
    toc: false
    grid: 
      margin-width: 10px
      body-width: 1000px
    resources: 
      - "*.js"
      - "*.mp3"
editor: source
---
<style>
body {
  background-color: #1D1D1B;
}
</style>

```{r echo=FALSE, message=FALSE, warning=FALSE}

library(tidyverse)
library(plotly)

app_df <- read.csv("C:/Users/Brayden Boyko/Downloads/322 Data CSV (2).csv")

df_long <- app_df %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "Years",
    values_to = "Value"
  ) %>%
  mutate(Years = as.numeric(str_remove(Years, "^X")))

colnames(df_long)[1] <- "Metric"

years <- sort(unique(df_long$Years))

line_df <- map_dfr(years, function(yr) {
  df_long %>%
    filter(Years <= yr) %>%
    mutate(frame = yr)
})

dot_df <- df_long %>%
  mutate(frame = Years)

p <- plot_ly()

p <- p %>%
  add_trace(
    data = line_df,
    x = ~Years,
    y = ~Value,
    frame = ~frame,
    color = ~Metric,
    type = 'scatter',
    mode = 'lines',
    line = list(width = 2),
    showlegend = TRUE
  )

p <- p %>%
  layout(
    title = "Financial Metrics Over Time",
    xaxis = list(title = "Year", color = "white"),
    yaxis = list(title = "Value (Millions)", color = "white"),
    plot_bgcolor = "#1D1D1B",
    paper_bgcolor = "#1D1D1B",
    font = list(color = "white")
  ) %>%
  animation_opts(frame = 600, easing = 'linear', redraw = FALSE) %>%
  animation_button(x = 1, xanchor = "right", y = 0, yanchor = "bottom") %>%
  animation_slider(
    currentvalue = list(prefix = "Year: ", font = list(color = 'white')),
    tickfont = list(color = 'white')
  )

p

```